---
title: "ENVS 193 - HW 5"
author: "Riley Zamora"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Introduction  


# Methods  

### Libraries  
```{r, message = FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
```


### Import data using here package  
```{r}
plant <- read.csv(here("hf109-01-sarracenia.csv")) %>% 
  clean_names() %>% 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```


## Part A  

## Part B  

## Part C  

### Visualize Missingness
```{r}
gg_miss_var(plant)
```

In this figure, we can see that chlorophyll, amass, sla, and num_lvs have missing data values. Chlorophyll and amass have over 10 missing, sla has 10 and num_phylls/num_lvs have less than 3 missing values. The rest of the variable have no missing data. 

### Remove NA  
```{r}
# creating a dataset without the missing values
plant_clean <- plant %>% 
  drop_na(sla, chlorophyll, amass, num_lvs, num_phylls)
```


## Part D  

### Visualize Correlation
```{r}
# calc Pearson's r for numerical values
plant_corr <- plant_clean %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")

# plot correlation values
corrplot(plant_corr,
         method = "ellipse",
         addCoef.col = "black")
```

The correlation figure illustrates the correlation values between various variables. A higher absolute value indicates a stronger correlation between the variables. For instance, the variables sla and amass exhibit the highest positive correlation, with a correlation coefficient of 0.32. Conversely, num_lvs and amass exhibit the strongest negative correlation, with a correlation coefficient of -0.31.  

## Part E  

```{r, message = FALSE}
plant_clean %>% 
  select(species:num_phylls) %>% 
  ggpairs() 
```

## Part F  

### Fitting Models
```{r}
null <- lm(totmass ~ 1, data = plant_clean) # just beta_0
full <- lm(totmass ~ species + feedlevel + # all predictors
             sla + chlorophyll + amass + 
             num_lvs + num_phylls, 
           data = plant_clean)
```

The null model only includes the intercept and response, `totmass`. The full model has all predictors in the dataset and the response as `totmass`.  

## Part G  
```{r}
par(mfrow = c(2, 2)) # facet the plots
plot(full) # will show diagnostic plots
```

```{r, warning = FALSE, message=FALSE}
# statistical diagnostics
check_normality(full) 
```
We get that "Non-normality of residuals detected (p < .001)." which means that we have to perform some transformations.

```{r, warning = FALSE, message = FALSE}
check_heteroscedasticity(full)
```
we get that "Heteroscedasticity (non-constant error variance) detected (p < .001)." more reason to perform a log transformation.  

# Part H  

### Fitting Log transform models  

I am going to perform a log transformation to the data to try and achieve normality and constant variance.  
```{r}
# creating log transformed models
null_log <- lm(log(totmass) ~ 1, data = plant_clean)
full_log <- lm(log(totmass) ~ species + feedlevel + 
                 sla + chlorophyll + amass + num_lvs + 
                 num_phylls, 
               data = plant_clean)
```


### diagnostic checking  

```{r}
check_normality(full_log) 
```
Our data is normal after the log transofrm.

```{r}
check_heteroscedasticity(full_log)
```
We have constant variance after the log transform! We will continue using the log model.

# Results  
















